What makes listeners infer teaching intentions?

### 2022-01-28 paper/manuscript.Rmd
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

## Apparatus
The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; https://www.psychopy.org/) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli (MIDI data) were played using the Mido library (1.2.9; https://mido.readthedocs.io/en/latest/) on a Max/MSP patcher (8.1.7; https://cycling74.com/products/max). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

## Stimuli
Piano recordings were used as stimuli. The stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. One piece of music was selected as a base piece and was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The piece was performed with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B) by multiple pianists. Therefore, all stimuli had the same musical sequence but had subtle differences in terms of musical expressions (e.g., timing, articulation and dynamics). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The stimuli were performed around 80 quarter-beats per minute (i.e., around 188 ms for 16th-note IOIs).

Ninety-six stimuli (MIDI recordings) were obtained from our previous experiments (Tominaga et al., *submitted*). In the previous experiments, we asked expert pianists to perform one piece of music with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). There were two experimental conditions. In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson). In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert). There were 453 valid performances (i.e., performances without any pitch error) for the teaching condition and 436 valid performances for the performing conditions. We randomly chose 24 articulation recordings and 24 dynamics recordings from the teaching condition as well as 24 articulation recordings and 24 dynamics recordings from the performing condition. It is important to note that the recordings from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

## Procedure
Upon arrival, participants read information sheet about the experiment and gave informed consent prior to participation. In the experiment, all instructions were displayed on a computer screen and an experimenter also explained the procedure. Participants were instructed that they were going to listen to piano recordings and would be asked if each recording was performed as if a pianist were teaching it to students (e.g., in a lesson / teaching) or as if a pianist were performing it to an audience (e.g., in a concert / performing). They responded by pressing either a yes (teaching / left arrow key) or no (performing / right arrow key) button. While listening to each stimulus, sheet music, which corresponded to the stimulus, was shown on a monitor in front of the participants. After making a judgment, the next recording was played (*Figure \ref{fig:procedure1}*).

There were two blocks and each block only contained recordings with one musical expressive technique of either articulation or dynamics. Each block consisted of 4 practice trials and 48 experimental trials. In each trial, participants listened to one recording only once. After listening the recording, they were asked to judge by pressing either the left arrow key (i.e., teaching) or the right arrow key (i.e., performing). Each recording was evaluated only once in the experiment. The order of the blocks was counterbalanced across the participants. The order of the recordings was randomised within each block.

>>>>

## Apparatus
The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; https://www.psychopy.org/) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido library (1.2.9; https://mido.readthedocs.io/en/latest/) on a Max/MSP patcher (8.1.7; https://cycling74.com/products/max). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

## Stimuli
We selected stimuli from our previous experiments (Tominaga et al., *submitted*; Experiment 1). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive techniques of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

In Tominaga et al., (*submitted*), participants were asked to perform the piece with either articulation or dynamics in two different conditions. In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson). In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert). In total, we obtained 453 valid performances (i.e., performances without any pitch error) from the teaching condition and 436 valid performances from the performing condition.

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al. (*submitted*). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

## Procedure
Upon arrival, participants read information sheet about the experiment and gave informed consent prior to participation. In the experiment, all instructions were displayed on a computer screen in front of the participants and an experimenter also explained the procedure. Participants were instructed that they were going to listen to piano recordings with musical expression (either articulation or dynamics), which were either produced as if a pianist were teaching the expression to students (e.g., in a lesson; teaching) or as if a pianist were performing it to an audience (e.g., in a concert; performing). In each trial, participants listened to one recording and were asked whether the recording was produced for teaching purposes or not. Participants responded by pressing either a yes (for teaching; left arrow key) or no (for performing; right arrow key) button. While listening to each recording, sheet music, which corresponded to the recording, was shown on the screen in front of the participants (*Figure \ref{fig:procedure1}*).

There were two blocks and each block only included recordings with one musical expressive technique of either articulation or dynamics. Each block consisted of 4 practice trials and 48 experimental trials. Each recording was evaluated only once in the experiment. The order of the blocks was counterbalanced across the participants. The order of the recordings was randomised within each block.

### 2022-01-28 paper/manuscript.tex
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; \url{https://www.psychopy.org/}) on a Mac Book Pro with Mac OS X 10.15.6. Stimuli (MIDI data) were played using the Mido library (1.2.9; \url{https://mido.readthedocs.io/en/latest/}) on a Max/MSP patcher (8.1.7; \url{https://cycling74.com/products/max}). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

Piano recordings obtained as MIDI data were used as Stimuli. Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. One piece of music was selected as a base piece, which was performed with a musical expressive technique of either articulation (\emph{Figure \ref{fig:stim1}}, A) or dynamics (\emph{Figure \ref{fig:stim1}}, B). The piece was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment. Therefore, all stimuli had the same musical sequence but had subtle differences in terms of musical expressions (e.g., timing, articulation and dynamics). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The stimuli were performed around 80 quarter-beats per minute (i.e., around 188 ms for 16th-note IOIs).

96 stimuli (MIDI recordings) were obtained from our previous experiments (Tominaga et al., \emph{submitted}). In the previous experiments, we asked expert pianists to perform one piece of music with a musical expressive technique of either articulation (\emph{Figure \ref{fig:stim1}}, A) or dynamics (\emph{Figure \ref{fig:stim1}}, B). There were two experimental conditions. In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson). In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert). There were 453 valid performances (i.e., performances without any pitch error) for the teaching condition and 436 valid performances for the performing conditions. We randomly chose 24 articulation recordings and 24 dynamics recordings from the teaching condition as well as 24 articulation recordings and 24 dynamics recordings from the performing condition. It is important to note that the recordings from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

@@

Upon arrival, participants read information sheet about the experiment and gave informed consent prior to participation. In the experiment, all instructions were displayed on a computer screen and an experimenter also explained the procedure. Participants were instructed that they were going to listen to piano recordings and would be asked if each recording was performed as if a pianist were teaching it to students (e.g., in a lesson / teaching) or as if a pianist were performing it to an audience (e.g., in a concert / performing). They responded by pressing either a yes (teaching / left arrow key) or no (performing / right arrow key) button. While listening to each stimulus, sheet music, which corresponded to the stimulus, was shown on a monitor in front of the participants. After making a judgment, the next recording was played (\emph{Figure \ref{fig:procedure1}}).

There were two blocks and each block only contained recordings with one musical expressive technique of either articulation or dynamics. Each block consisted of 4 practice trials and 48 experimental trials. In each trial, participants listened to one recording only once. After listening the recording, they were asked to judge by pressing either the left arrow key (i.e., teaching) or the right arrow key (i.e., performing). Each recording was evaluated only once in the experiment. The order of the blocks was counterbalanced across the participants. The order of the recordings was randomised within each block.

>>>>

The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; \url{https://www.psychopy.org/}) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido library (1.2.9; \url{https://mido.readthedocs.io/en/latest/}) on a Max/MSP patcher (8.1.7; \url{https://cycling74.com/products/max}). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

We selected stimuli from our previous experiments (Tominaga et al., \emph{submitted}; Experiment 1). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive techniques of either articulation (\emph{Figure \ref{fig:stim1}}, A) or dynamics (\emph{Figure \ref{fig:stim1}}, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

In Tominaga et al., (\emph{submitted}), participants were asked to perform the piece with either articulation or dynamics in two different conditions. In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson). In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert). In total, we obtained 453 valid performances (i.e., performances without any pitch error) from the teaching condition and 436 valid performances from the performing condition.

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al.~(\emph{submitted}). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

@@

Upon arrival, participants read information sheet about the experiment and gave informed consent prior to participation. In the experiment, all instructions were displayed on a computer screen in front of the participants and an experimenter also explained the procedure. Participants were instructed that they were going to listen to piano recordings with musical expression (either articulation or dynamics), which were either produced as if a pianist were teaching the expression to students (e.g., in a lesson; teaching) or as if a pianist were performing it to an audience (e.g., in a concert; performing). In each trial, participants listened to one recording and were asked whether the recording was produced for teaching purposes or not. Participants responded by pressing either a yes (for teaching; left arrow key) or no (for performing; right arrow key) button. While listening to each recording, sheet music, which corresponded to the recording, was shown on the screen in front of the participants (\emph{Figure \ref{fig:procedure1}}).

There were two blocks and each block only included recordings with one musical expressive technique of either articulation or dynamics. Each block consisted of 4 practice trials and 48 experimental trials. Each recording was evaluated only once in the experiment. The order of the blocks was counterbalanced across the participants. The order of the recordings was randomised within each block.
